\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-15 arch.tex
Added an illustration based on Fig. 14 to explain how the existing plan supports synchronization search

  each reactor declares its planning look-ahead and expected latency
  \comment{$\pi \& \lambda$ both defined here for the first time}
  before being able to produce plan to the agent depending on where it
  is located in a reactor dependency graph such as shown in
  Fig. \ref{fig:agent}.

@@

We introduce this requirement as a new type of flaw for the \eu
framework defined as follows:

@@

\end{enumerate}
\end{proposition}

Intuitively the first choice enforces an intertail value assumption
where the reactor will tend to maintain its current as long as the
model and external observations allow it. The second choice favorise
the adavnce of the reactor current plan if any; this direct the {\em
  synchronization} decision toward assuming that its current plan is
properly executed. Finally if none of these two choices did provide a
valid solution -- meaning that both lead to an inconsittant plan --
the reactor tries all the possible state values remaining for this
timeline until it finds one that allows it to identify its state as
required by \rx architecture.

@@

\item Resolve the current state value of each internal timelines based
  on Proposition \ref{prop:csf:resolve}.
\end{enumerate}

Steps $2$ \& $3$ are handled by the \eu solver from algorithm
\ref{alg:europa:solve}. We only enforce that the {\em current state
  flaws} are selected last by the algorithm in order to follow the
sequencing given above. As {\em synchronization} focusses on the
execution frontier wee also filter out all flaws that are outside of
the current tick scope. This mean that if this flaw relates to a token
that is {\em necessarily} ending before $\tau$ or starting after
$\tau+1$ then the solver will ignore it. This helps to reduce the
scope of the synchronization to exclusively the current state of the
reactor and reduce dramatically the search space for the {\em
  synchronization} solver. In order to do so we need to make the
assumption that the current state value of an {\em internal} timeline
does not depend on the future or more accurately that any choices made
during this synchronization will not lead to a future
inconsistency. Such an assumption impacts the set of possible domains
the reactor can support while remaining complete. Take for example
this rule from our \texttt{Shopping} model:

@@

We do so by relaxing the constraint in line $5$ from a strict
\texttt{equals} to a \texttt{contains} and to add the new constraint
at line $11$ which ensures that the last \texttt{Going} ends up in the
location the agent actually wants to be at. This first level of
relaxation will therefore allow the agent to use multiple
\texttt{Going} tokens should the first one fail to reach the
objective.  % The problem can even be relaxed further if needed but
% this relaxation gives already more felxibility for the reactor to
% recover from local failures.
As the simple example above shows, synchronization involves a
deliberation process with the risk that the search for a solution is
not a straightforward sequence of unit decisions. Since such a process
is required by the architecture at \emph{every} tick for \emph{every}
reactor the model needs to be carefully designed.

At the engine level careful choices were made in both the focus of the
search -- limited only to tokens that can potentially overlap the
current tick -- and heuristics where the internal state flaw is
considered last during the search process. The later limits the amount
of backtracking for a solution -- or at least how deep one needs to
backtrack in a decision tree. Similarly the sequence in which we
evaluate the possible resolution for such a flaw -- presented
previously -- was selected according to some simple assumptions that % takes
% benefit of a possibly assume the existing plan by assuming that 
if the outcome of synchronization appears consistent with our current
plan then it probably reflects that the plan has executed
\comment{that the past is monotonic?}. Such an assumption helps direct
the internal state evaluation and in nominal cases, proves to be a
sound solution.

One final aspect to take into account is that from a \rx point of
view, \emph{synchronization is the only critical task of a
  reactor}. Such a process is used in order for the agent to identify
its current state and maintain it consistently throughout the agent's
lifetime. For a reactor not being able to find a consistent solution
during synchronization -- implying that after evaluating all possible
solutions provided by a model it was unable to find a consistent plan
-- not only indicates that a reactor is unable to explain the state of
the world, but more importantly it may jeopardize other reactors that
rely on its {\em internal} state. For this reason a failure to
synchronize immediately results in the agent killing a reactor and
notifying any reactor that depends on its {\em internal} state
variables; it does so by posting the special observation
\textsf{Failed}. And by doing so we allow for graceful degradation
where other reactors can readapt their plans as they receive this
observation.

@@

The most obvious case is how synchronization allows propagate of {\em
  external} observations when a new tick occurs. In our presentation
of synchronization in Section \ref{sec:arch:synch}, we isolated the
problem by assuming that the reactor has no future goal or more
accurateely no plan to enforce.  By doing so we were able to develop
the way we augment the \eu solver to be able to propagete {\em
  external} observations in order to identify the reactor {\em
  internal} state for the same tick. By doing so we were able to focus
on synchronization as a state evaluation process that can be resolved
within the \eu framework with few extensions. Similarly we presented
deliberation as an abstraction while not detailing how the disruptive
nature of synchronization could impact the way planning could evolve.

@@

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.7\columnwidth]{figs/shoping_exec_t0}
  \caption{\small Shopping example: A plan produced before
    synchronization.  Tokens in red indicate observations, tokens in
    blue the goal that this plan attempts to solve for. Arrows
    indicate temporal constraints.}
  \label{fig:shop:exec0}
\end{figure}

Consider that at the next tick $\tau' = \tau+1$ the reactor managing
agent location is able to change the location state variable to
\texttt{Going}. \rx informs our reactor of this new observation which
is added in the plan structure as a fact -- starting exactly at
$\tau'$ and lasting for an unknown duration which is at least 1
tick. As this new observation is integrated in the plan through
synchronization, the solver can identify the similarity between this
new observation and the next token in the existing plan. This implies
that instead of initiating an exhaustive search of all possible
implications, the synchronization process starts its execution with
the assumption that these new observations are consitent with the
currently maintained plan. As a consequence, at the end of
synchronization we will likely have the plan depicted in
Fig. \ref{fig:shop:exec1}. Which just propagated the information
provided by the new observation in the exisiting plan (restricting in
turn the start time of \texttt{Going} and by propagation of the
constraints \texttt{Go} to be {\em exactly} $\tau'$). While focusing
uniquely on the execution frontier ($\tau'$ in
Fig. \ref{fig:shop:exec1}), the \eu solver took benefit of the
exisiting plan to evaluate a solution that asumes that the observation
is consistent with what was decided inside this reactor.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.7\columnwidth]{figs/shoping_exec_t1}
  \caption{Shopping example : Result of synchronization after the {\em
      Going} observation was received by the reactor with the plan
    showed in Fig. \ref{fig:shop:exec0}.}
  \label{fig:shop:exec1}
\end{figure}

This choice is directed not only by some natural heurisitc choices
made in the solver but even more enforced in general by the heuristic
order of the decision point reolution choices for our {\em internla}
state flaw presented in section \ref{sec:arch:sync} {\em\color{red}
  TODO: need to rework the presentation of thee choices so I can refer
  to them more directly -- algorithm or pseudo theorem should do the
  trick}. Should such a flaw be revealed during synchronization --
meaning that the current state of an {\em internal} state varaibale is
not fully grounded -- the sequencing of the choices will atempt first
a conservative choice in term of maintaining previous state, then an
optimisitic choice in term of the execution/advance in our current
plan to finally attempt other choices that were not predicted by the
model. The two first possible choices are strongly directed by the
exisiting plan which help direct the search within this frame. This
often allow to have a synchronization that is much more directed and, 
in nominal cases such as the one we showed resolve the deliberation 
in few steps (often without any backtrack in the search).

The synchronization process is also in this process supporting fully
the tracking of the plan execution as it specifies tokens parameters
such as the time points and allow this to propagate through the plan
in order to idenitfy if the plan remains valid -- as in our example --
or that it cannot be executed as it is. The later is often identified
by the synchronization failing to find a consistent solution. for
example dshould the \texttt{Going} observation we receive be to the
\texttt{hardwareStore}, this would break the current plan of our
reactor as we cannot go simultaneously at 2 different places as
illustrated in Fig. \ref{fig:shop:relax}-1. as synchronization is not
possible in this context our strategy is to fully relax all decisions
made by the previous plan steps by just keeping n the plan all the
past observations in both {\em external} and {\em internal} state
variables and the exsiting {\em internal} goals received by this
reactor (\texttt{Own(Milk)} in our example) resulting on the plrtial
plan sghowned in Fig. \ref{fig:shop:relax}-2. In this process the
reactor inform the agent that all {\em external} goals it requested
for the future are note valid anymore which often result in the ownere
of htese timelines to not have to maintain this goal anynmore.

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.7\columnwidth]{figs/shoping_exec_relax}  
  \caption{Illustration of a conflict during synchronization and the
    resulting relaxed plan for recovery} 
  \label{fig:shop:relax}
\end{figure}

As the plan is not anymore in the way the synchronization can resume
and find a solution as we initial lly described. The newly pending
goals will enforce at the next \texttt{step} to resumed deliberation
on this new fully relaxed partial plan. By the simple use of
synchronization the reactor was not only able to identify that its
plan was not executable anymore but recover from it while offering a
new incomplete partial plan to be resolved during the next
deliberation \texttt{steps}. This on its own ogfer the core
functionality one should except to close the loop between planning and
execution in the sense that synchronization plays the role of what is
typically named as the {\em Executive} in other architectures
{\em\color{red} references to diverse archis, 3-tiered, laas, claraty,
  B Wiliams, and many more}.

In our case though the interaction between planning and execution does
not stop at this level. A core point of the overal schedulling of both
{\em synchronization} and {\em planning} steps resides on our
architecture enforsincing that the synchroniuzation process can occur
in between any step of the plan. This lead us to the fact that not
only synchronization is imnfluenced by the outcome of planning but
this influence goes also in the alternate direction. Specifically when
synchronuiation occuers before we found a complete plan, it will
inject as we described new facts (the {\em external} observations and
resulting {\em internal} state) in the plan database. This
informatiuon will be present when the next planning \texttt{step} will
resume. Therefore they are not intergal part of the plan resolution
(generating potentially new flaws for the planner or contributing to
resolve previously exisiting flaws). The planning process is therefore
perturbed internallyand fully informed on how the world evolves as it
is planning.  This aspect trueluy contrast to other approaches we have
identified in the literature which avoid to perturb the planner during
it search as it is not compatible with the ``off-line planning''
assumption.

By our design choices -- both at the architecture level and the
integration of the \eu framework for embedded deliberation -- we
were able to implement an architecture that avoid this limitation
while remaining in a formal frame that still allow for deliberation
and model-based agent control. We believe that this tighter
integration of planning and execution allow the system to ba more
acute of its environment allowing better informaed decision and
tighter reaction. This is possible while still integrating information
such as a long latency for the reactors as long as one can ensure that
synchronization of al the reactors and deliberation steps can be done
in a time that is resonably small in comparison of the agent tick
duration.

>>>>

  each reactor declares its planning look-ahead $\pi$ and expected
  latency $\lambda$
  % \comment{$\pi \& \lambda$ both defined here for the first time}
  before being able to produce plan to the agent depending on where it
  is located in a reactor dependency graph such as shown in
  Fig. \ref{fig:agent}.

@@

We introduce this requirement as a new type of flaw for the \eu
framework defined as follow:

@@

\end{enumerate}
\end{proposition}
Intuitively the first choice enforces an intertial value assumption
where the reactor will tend to maintain its current as long as the
model and external observations allow it. The second choice favorise
the advance of the reactor current plan if any; this direct the {\em
  synchronization} decision toward assuming that its current plan is
properly executed. Finally if none of these two choices did provide a
valid solution -- meaning that both lead to an inconsistant plan --
the reactor tries all the possible state values remaining for this
timeline until it finds one that allows it to identify its state as
required by \rx architecture.

@@

\begin{enumerate}
\item Integrate the external state as provided by the owner of each
  external timeline into the plan database.
\item Propagate this information in the plan database following the
  model $\mathcal{M}$ of this reactor.
\item Resolve the current state value of each internal timelines based
  on Proposition \ref{prop:csf:resolve}.
\end{enumerate}

%%%% Extra paragraph:  targeted to a more AI centric audience
% The step 1 translates new {\em external} observations received from
% \rx into europa tokens with a start time restricted to $\tau$ or, for
% the timelines that did not receive a new observation,  restrict the
% previous token end time to the interval $[\tau+1, \infty)$. All of
% these tokens are produced as facts and consequently cannot be 
% excluded from the plan. They are also maintained by the reactors as
% observation tokens which will allow to avoid to mistakenly remove 
% them when one of the two reactor solvers attempt to relax former
% decisions made.\comment{This relaxation is developed on the last 
%   subsection further in depth}
%%%%

Steps $2$ \& $3$ are handled by the \eu solver from algorithm
\ref{alg:europa:solve}. We only enforce that the {\em current state
  flaws} are selected last by the algorithm in order to follow the
sequencing given above. As {\em synchronization} focusses on the
execution frontier wee also filter out all flaws that are outside of
the current tick scope. This mean that if this flaw relates to a token
that is {\em necessarily} ending before $\tau$ or starting after
$\tau+1$ then the solver will ignore it. This helps to reduce the
scope of the synchronization to exclusively the current state of the
reactor and reduce dramatically the search space for the {\em
  synchronization} solver. In order to do so we need to assume 
that the current state value of an {\em internal} timeline does not 
depend on the future or more accurately that any choices made during 
this {\em synchronization} will not lead to a future inconsistency. 
Such an assumption impacts the set of possible domains the reactor 
can support while remaining complete. Take for  example this rule from 
our \texttt{Shopping} model :

@@

We do so by relaxing the constraint in line $5$ from a strict
\texttt{equals} to a \texttt{contains} and to add the new constraint
at line $11$ which ensures that the last \texttt{Going} ends up in the
location the agent actually wants to be at. This first level of
relaxation will therefore allow the agent to use multiple
\texttt{Going} tokens should the first one fail to reach the
objective.  % The problem can even be relaxed further if needed but
% this relaxation gives already more flexibility for the reactor to
% recover from local failures.
As the simple example above shows, synchronization involves a
deliberation process with the risk that the search for a solution is
not a straightforward sequence of unit decisions. Since such a process
is required by the architecture at \emph{every} tick for \emph{every}
reactor the model needs to be carefully designed.

One critical aspect to take into account is that from a \rx point of
view, synchronization is the only critical task of a reactor. Such a
process is used in order for the agent to identify its current state
and maintain it consistently throughout the agent's lifetime. For a
reactor not being able to find a consistent solution during
synchronization -- implying that after evaluating all possible
solutions provided by a model it was unable to find a consistent plan
-- not only indicates that a reactor is unable to explain the state of
the world, but more importantly it may jeopardize other reactors that
rely on its {\em internal} state. For this reason a failure to
synchronize immediately results in the agent killing a reactor and
notifying any reactor that depends on its {\em internal} state
variables; it does so by posting the special observation
\textsf{Failed}. And by doing so we allow for graceful degradation
where other reactors can readapt their plans as they receive this
observation.

Finally, at the engine level careful choices were made in 
both the focus of the search -- limited only to tokens that 
can potentially overlap the current tick -- and heuristics where 
the {\em current state flaw} is considered last during the search
process. The later limits the amount of backtracking for a solution 
-- or at least how deep one needs to backtrack in a decision tree. 
Similarly the sequence in which we evaluate the possible resolution 
for such a flaw -- presented in Proposition \ref{prop:csf:resolve} -- 
was selected according to  some simple assumptions that any external 
observation that is compatible with an existing token in the current
plan reflects the execution of the plan itself. In other words this
means that our solver always prefer to use tokens already existing in
the plan in order to identify its current internal state. By doing so
we directs the synchronization search by exploiting the structure of
the plan produced during previous deliberation steps. This implies
that a deliberation steps have been done and did impact the plan
structure and calls for a description on how our \eu reactor resolve
such steps which we are now going to describe further.

@@

The most obvious case is how synchronization allow propagate of {\em
  external} observations as a new tick occurs. In our presentation of 
{\em synchronization}, we isolated the problem by stating
that we will assume at this point that the reactor has no future goal
or more accurately no plan to enforce. By doing so we
were able to develop the way we augment the \eu solver to be able
to propagate {\em external} observations in order to identify the
reactor {\em internal} state for the same tick. This allowed us 
to focus on synchronization as a state evaluation process that can
be resolved within the \eu framework with few extensions. Similarly
we presented deliberation at this fairly high level not necessarily
discussing how the disruptive nature of synchronization in this
process could impact the way this planning could evolve.

@@

\begin{figure}[!htb]
  \centering
  \subfloat[\small A plan produced before synchronization.]{\label{fig:shop:exec0}\includegraphics[width=0.45\columnwidth]{figs/shoping_exec_t0}}
  \subfloat[\small Result of synchronization after the {\em Going}
  observation was received by the reactor with the plan showed in
    \ref{fig:shop:exec0}.]{\label{fig:shop:exec1}\includegraphics[width=0.48\columnwidth]{figs/shoping_exec_t1}}
    \caption{\small Shopping example: Tokens in red indicate
      observations, tokens in blue the goal that this plan attempts to
      solve for. Arrows indicate temporal constraints. }
  \label{fig:shop:exec}
\end{figure}

Consider that at the next tick $\tau' = \tau+1$ the reactor managing
agent location is able to change the location state variable to
\texttt{Going}. \rx informs our reactor of this new observation which
is added in the plan structure as a fact -- starting exactly at
$\tau'$ and lasting for an unknown duration which is at least 1
tick. As this new observation is integrated in the plan through
synchronization, the solver can identify the similarity between this
observation and the next token in the existing plan. Instead of
initiating an exhaustive search of all possible implications, the
synchronization process starts its execution with the assumption that
these new observations are consistent with the currently maintained
plan. As a consequence, at the end of synchronization the plan
depicted in Fig. \ref{fig:shop:exec1} is likely. This is a result of
propagation of information provided by the new observation in the
existing plan while restricting the start time of \texttt{Going} and
propagating temporal constraints, making \texttt{Go} to start {\em
  exactly} at $\tau'$. In doing so, the \eu solver's focus on the
execution frontier ($\tau'$ in Fig. \ref{fig:shop:exec1}) takes
advantage of the existing plan to evaluate a solution that assumes
that the observation is consistent with what was decided by the
reactor.

% \begin{figure}[!htb]
%   \centering
%   \includegraphics[width=0.7\columnwidth]{figs/shoping_exec_t1}
%   \caption{Shopping example : Result of synchronization after the {\em
%       Going} observation was received by the reactor with the plan
%     showed in Fig. \ref{fig:shop:exec0}.}
%   \label{fig:shop:exec1}
% \end{figure}

This choice is directed not only by heuristic choices made in the
solver but also enforced by the heuristic order of the decision point
resolution choices for the {\em current  state flaw} presented in
Proposition \ref{prop:csf:resolve}. Indeed, as we stated then the two
first choices tend to privilege a conservative view in respect to the
existing plan. It attempts first to extend the previous observation we
produced for this {\em internal} timeline in order to avoid to commit
to early in the plan advance. If the formaer failed it will then
attempt to start the next exisiting token in the plan which alow the
reactor to advance its plan execution. Both of these two initial
choices atetempt to resolve the plan with the least amount of
perturbation of the plan. They just constrain further the start and
end time of the exisiting tokens which were provided by the previous 
deliberations steps. This means that in a nominal situation where the
plan this reactor produced is properly executed, the synchronization
remains limited to restrict time points in the plan and progate their
impact. Most of the heavy dutty work of creating new tokens in the
plan and schedule them -- both being much more difficult problems 
to resolve -- have been managed during deliberation steps. In that
sense we see that most of the time synchronization is limited to very
few node in \eu search space with often no backtrack required. 
Fig. \ref{fig:shop:exec} shows how the existing plan can support the
synchronization process. In that case, the observation \texttt{Going}
occuring at $\tau'$ was fully compatible with the one already present
in our plan, by updating the plan and progating the equality
constraints, the resolution of the {\em current state flaw} for the
{\em agent} timeline became obvious as the token \texttt{Go} from 
the plan presented itself as a natural solution.

The synchronization may become slightly nore costfull when we need to
identify a new token to be injected by the plan (the 3rd choice in
Proposition \ref{prop:csf:resolve}) as it needs to explore on tis own
all the possible token types this timeline accept as value. This is
the reason this choice is evaluated last in our flaw resolution in
order to avoid such complex action during synchronization unless 
necessary. In that sense,  and thanks to the carefull choice on our
synchronization heurisitic we see that the plan produced during \rx
deliberation steps helps sthe synchronization process to be more
efficient.

% Should such a flaw be revealed 
% during synchronization -- such that the current state of
% an {\em internal} state variable is not fully grounded -- the
% sequencing of the choices will attempt first a conservative choice in
% maintaining previous state and only then try a more optimistic
% approach in terms of the \comment{I don't follow this last bit. Is
%   this token addition that you're talking about?} execution/advance in
% our current plan to finally attempt other choices that were not
% predicted by the model. The two first possible choices are strongly
% directed by the existing plan which helps direct the search within
% this frame. This often allow to have a synchronization that is much
% more directed and, in nominal cases such as the one we showed resolve
% the deliberation in few steps (often without any backtrack in the
% search) \comment{All this is confusing; probably reword}.

Synchronization also supports the tracking of plan execution. It does
so by specifying token parameters such as time points that are
propagated through the plan in order to identify if the plan remains
valid -- as in our example -- or that it cannot be executed as it
is. The later is often identified by the synchronization failing to
find a consistent solution. For example should the \texttt{Going}
observation received be to the \texttt{hardwareStore}, this would
break the current partial plan in the reactor as the robot cannot
possible be going simultaneously to $2$ different locations; see
Fig. \ref{fig:shop:relax1}. As synchronization is not possible in such
a situation the strategy followed is to relax all decisions made by
previous plan steps. This is done by keeping all past observations in
both {\em external} and {\em internal} state variables as well as the
existing {\em internal} goals (\texttt{Own(Milk)} in this example)
received by this reactor. The resulting partial plan is shown in
Fig. \ref{fig:shop:relax2}. In this process the reactor informs the
agent that all {\em external} goals it requested for the future are no
longer valid which results in turn, in timeline owners not having to
maintain such goals.

\begin{figure}[!htbp]
  \centering
  \subfloat[Conflicting Observation]{\label{fig:shop:relax1}\includegraphics[width=0.7\columnwidth]{figs/shoping_exec_relax-1}}\\
  \subfloat[Relaxed Partial Plan]{\label{fig:shop:relax2}\includegraphics[width=0.7\columnwidth]{figs/shoping_exec_relax-2}}
  \caption{\small Illustration of a conflict during synchronization
    and the resulting relaxed plan for recovery.}
\end{figure}

As the plan is no longer conflicted, synchronization can resume and
find a solution described above. Newly pending goals would be enforced
at the next \texttt{step} to resume deliberation on this relaxed
partial plan. By using synchronization the reactor was not only able
to identify that its plan was not consistent but able to recover from
it by offering a new incomplete partial plan to be resolved in the
next deliberation \texttt{steps}. % This on its own offer the core
% functionality one should except to close the loop between planning and
% execution in the sense that 
Synchronization plays the role of what is typically termed as the {\em
  Executive} in other architectures \cite{gat98, alami:1998p820,
  mus98, williams03, Nesnas:2003do}.

In \rx however the interaction between planning and execution does not
stop at this level. The overall scheduling of both {\em
  synchronization} and {\em planning} is such that the synchronization
can occur in between any planning step. Not only is synchronization
influenced by the outcome, but it also influences
planning. Specifically when synchronization occurs before a complete
plan is found, it will inject new facts ({\em external} observations
and resulting {\em internal} state) in the plan database available for
inference when the next planning \texttt{step} resumes. On this next
step the \eu solver used for planning will have these new facts
included in the partial plan which can introduce new flaws in the plan
while potetnially having reolved previously exisiting flaw. The
deliberation process is therefore perturbed externally by the {\em
  synchronization} and consequently fully informed of the world state
evolution as it is planning.  This specific aspect contrasts with
other more classical approaches we have seen, which avoid 
perturbing the planner during its search as it is not compatible 
with the ``off-line planning'' assumption.

By our design choices -- both at the architecture level and the
integration of the \eu framework for embedded deliberation -- we are
able to implement an architecture that avoids this critical off-line
planning limitation while remaining within a formal framework that
allows for deliberation and model-based agent control. We believe that
this tight integration of planning and execution allows the system to
be more acute vis-a-vis the environment allowing informed
decision-making and a tighter reaction loop. This is possible while
integrating information despite a long latency for the reactors, as
long as one can ensure that synchronization of all reactors and
deliberation steps can be done within a time that is resonably small
in comparison to the agent tick duration.
